\section{Introduction}

Recent years have seen the gig economy reshape how services are delivered in urban environments, particularly in the context of food and package transportation. Mobile platforms have made it possible to coordinate short-term delivery work efficiently at scale, creating new possibilities—and new challenges—for digital service design.

This project explores one such possibility through the development of \textbf{LiftDrop}, a mobile application designed to simulate a real-time courier coordination system. The focus is on delivering a technically sound, user-aware platform that supports dynamic task assignment and live communication between clients and couriers.



\vspace{3mm}

\subsection{Motivation}

This project was inspired by the increasing influence of the gig economy on how urban delivery services are structured and experienced. With food and package delivery systems becoming an everyday part of modern life, we were interested in exploring the technical challenges involved in building a platform that simulates such systems—particularly one that emphasizes responsive behavior, performance-based decision-making, and real-time coordination.

\vspace{3mm}

For this purpose, the project proposes a solution that offers a clear and effective order assignment mechanism based on courier performance and proximity. It should also support seamless order cancellation by couriers—both before and after picking up the delivery—as well as intelligent order creation for clients, which automatically selects the nearest branch of a restaurant chain when a specific location is not provided.

\subsection{Organization}

The remainder of this report is organized as follows:
\begin{itemize}
    \item \textbf{Background and Project Scope}: Overview of gig economy delivery platforms and relevant technologies.
    \item \textbf{System Design}: A comprehensive view of the platform’s design, including the overall system architecture, user interface structure, and client/courier journey flows.
    \item \textbf{Technical Implementation Details}: Description of technologies, development tools, and API interactions used during the project.
\end{itemize}
